Quantum Magic Games

### 2023-07-06 paper/main.tex
Paper => typos
refactor intro

In this project, we will explore the Mermin-Peres magic square game
(hereinafter referred simply as magic square). The origins of this
game date back to the ninities. It was first described in the works of
Mermin~\cite{mermin:1990, mermin:1993} and of
Peres~\cite{peres:1990, peres:1997}. Their results provided (as per the title of
Mermin's review letter) a simplied proof for the Kochen-Specker
theorem~\cite{kochen:1967}. Later, in 2002, Aravind demonstrated how
to transform Mermin-Peres's proof of the Kochen-Specker theorem into a
proof of Bell's theorem~\cite{aravind:2002}. Bell's
theorem~\cite{bell:1964} shows the incompatibility of hidden variables
(i.e. determinism) and locality, and Kochen-Specker's theorem
demonstrates the conflicting nature of determinism and
noncontextuality (locality can be regarded as a special case of
contextuality). These sort of results may seem
challenging, but their descriptions can be greatly simplified by
modeling them as quantum games.  In particular, it is extremely easy
to show that there cannot be a classical solution to the the magic
square game, and to convince an observer that something ``magical''
(classically impossible) is happening in a successful implementation
of a quantum winnin strategy~\cite{brassard:2005}.

\paragraph{Outline of this paper.}
In Section~\ref{sec:magic-square}, we detail the intrinsics of the
magic square game, classical approaches and why it is impossible to
systematically win the game classically, how and why the quantum
strategy works, and how to implement it. We also briefly discuss how
the game translates to the proof of incompatibility of a
noncontextual hidden-variable theory. Then, in
Section~\ref{sec:extending}, we describe a generalization of games
that follow the same style of the magic square game, given
by Arkhipov~\cite{arkhipov:2012}. Finally, Section~\ref{sec:conc} concludes this report.

@@

\subsection{Quantum Solution}
\label{sec:quantum-sol}
The quantum strategy for the magic square game consists of Alice and
Bob carrying entangled qubits with them. Then, when Charlie asks Alice
(respectively Bob) to fill the entries of the \(i\)-th row
(respectively \(j\)-th column), Alice measures her two qubits three
times. The outcome of each measurement gives the value for each cell
(in practice, they do not even need the third measurement, since they
can determine the third value from the first two). Key to this
strategy is that the outcomes of all possible measurements are
\(\pm 1\). The initial, entangled state is (subscripts \(A\) and \(B\)
indicate, respectively, Alice's and Bob's qubits)
\begin{equation*}
  \ket{\psi} = \frac{1}{2}\left(
    \ket{00}_{A} \otimes \ket{00}_{B} + \ket{01}_{A} \otimes \ket{01}_{B} +
    \ket{10}_{A} \otimes \ket{10}_{B} + \ket{11}_{A} \otimes \ket{11}_{B}
  \right).
\end{equation*}
Figure~\ref{fig:circuit-psi} shows the quantum circuit that can be
used to prepare the initial state \(\ket{\psi}\).

>>>>

In this project, we will explore the Mermin-Peres magic square game.
% (hereinafter referred simply as magic square).
The origins of this
game date back to the ninities. It was first described by
Mermin~\cite{mermin:1990, mermin:1993} and 
Peres~\cite{peres:1990, peres:1997}. Their results provided
% (as per the title of Mermin's review letter)
a simplied proof for the Kochen-Specker
theorem~\cite{kochen:1967}. Later, in 2002, Aravind demonstrated how
to transform Mermin-Peres's proof of the Kochen-Specker theorem into a
proof of Bell's theorem~\cite{aravind:2002}. Bell's
theorem~\cite{bell:1964} shows the incompatibility of hidden variables
(i.e. determinism) and locality, and Kochen-Specker's theorem
demonstrates the conflicting nature of determinism and
noncontextuality (locality can be regarded as a special case of
contextuality). These sort of results may seem
challenging, but their descriptions can be greatly simplified by
modeling them as quantum games.  In particular, it is easy
to show that there cannot be a classical solution to the the magic
square game, and to convince an observer that something ``magical''
(classically impossible) is happening in a successful implementation
of a quantum winning strategy~\cite{brassard:2005}.

\paragraph{Outline of this paper.}
In Section~\ref{sec:magic-square}, we detail the magic square game,
classical approaches and why it is impossible to systematically win
the game classically, how and why the quantum strategy works, and how
to implement it. We also briefly discuss how the game translates to
the proof of incompatibility of a noncontextual hidden-variable
theory. Then, in Section~\ref{sec:extending}, we describe a
generalization of games of the same style, given by
Arkhipov~\cite{arkhipov:2012}. Section~\ref{sec:conc} concludes this
report.

@@

\subsection{Quantum Solution}
\label{sec:quantum-sol}
The quantum strategy for the magic square game consists of Alice and
Bob carrying entangled qubits with them. Then, when Charlie asks Alice
(respectively Bob) to fill the entries of the \(i\)-th row
(respectively \(j\)-th column), Alice measures her two qubits three
times. The outcome of each measurement gives the value for each cell
(in practice, they do not even need the third measurement, since they
can determine the third value from the first two). Key to this
strategy is that the outcomes of all possible measurements are
\(\pm 1\). The initial entangled state is (subscripts \(A\) and \(B\)
indicate, respectively, Alice's and Bob's qubits)
\begin{equation*}
  \ket{\psi} = \frac{1}{2}\left(
    \ket{00}_{A} \otimes \ket{00}_{B} + \ket{01}_{A} \otimes \ket{01}_{B} +
    \ket{10}_{A} \otimes \ket{10}_{B} + \ket{11}_{A} \otimes \ket{11}_{B}
  \right).
\end{equation*}
Figure~\ref{fig:circuit-psi} shows the quantum circuit that can be
used to prepare the initial state \(\ket{\psi}\).
